\begin{abstract}
Courses are experimental interventions in robot racing: their geometry, topology,
semantics, and sampling distribution condition what a controller learns and what an
evaluation can establish. Yet tracks, gate sequences, waypoint routes, and buoy
courses are represented and generated through fragmented conventions spanning ground,
aerial, and maritime robotics, procedural content generation, scenario generation,
and simulation. This survey treats the robot course as the common unit of analysis.
It develops a vocabulary that separates course objects, representations,
realizations, generators, and artifact-management operations; synthesizes generation
methods and domain-dependent feasibility constraints; and proposes auditable metrics
for validity, geometry, continuous multidimensional task spectra, feasible diversity,
reproducibility, and simulator interoperability. Building on this synthesis, we define
a prospective benchmark protocol with immutable course identities, explicit raw and
repaired outcomes, paired policy evaluation, and separate training-distribution and
held-out-suite requirements. We also specify a portable \texttt{CourseSpec} contract,
simulator and reinforcement-learning adapter requirements, and an agenda for open
reference implementations. These contracts and corpora are proposals, not released
capabilities. The technical synthesis is qualitative and source-backed; quantitative
taxonomy and prevalence claims are withheld because two independent contribution-
coding assessments did not meet the prespecified all-field reliability gate. The
result is a framework for making course generation itself inspectable, comparable,
and reusable across robot-racing domains without reducing difficulty to a single
label or conflating generator quality with controller performance.
\end{abstract}
