\section{Domain Constraints}
\label{sec:domain-constraints}

A course becomes operational only when its spatial structure is connected to
the traversal semantics and feasibility conditions of a domain.  The same
polyline, gate set, or waypoint list can be adequate in one setting and
underspecified in another.  This section therefore treats constraints as part of
the course contract, not as decorative detail added after generation.  The
comparison preserves domain-specific requirements while identifying a common
abstraction suitable for portable course descriptions.  Competition rules and
standards are used only as supporting evidence for task semantics, coordinates,
or realization requirements; they are not treated as generators.

\subsection{Ground Courses}

Ground courses are usually constrained by a route or corridor embedded in a
surface-bound world.  At the topological level, the course must state whether it
is an open route, a closed circuit, a branching road network, or a route through
a network.  A road graph alone is insufficient for an evaluation task when it
does not identify admissible successors, start and terminal conditions, or the
lane-level relation to be traversed.  Constructive road systems make this
dependence visible: PGDrive joins road blocks through declared sockets and
configures lanes, width, and curvature as course-defining parameters
\cite{Li2020ImprovingGeneralization}; JunctionArt constructs intersection
connectivity and lane geometry from control lines and road-design rules
\cite{Muktadir2023ProceduralGeneration}.

The geometric contract needs a centerline or boundary representation together
with width, lane, and boundary semantics.  A centerline without a width profile
cannot state lateral clearance; paired boundaries without a successor relation
cannot state intended progress.  Curvature and curvature rate constrain the
ability of a vehicle to remain within the corridor under a declared speed and
control envelope.  At intersections, turn geometry, field of view, conflict
area, and the number of possible connecting lanes can all change the sensing and
planning task even when the overall network topology is unchanged
\cite{Muktadir2023ProceduralGeneration}.  Such descriptors should be recorded
in physical units and evaluated against a stated vehicle envelope, rather than
collapsed into a generic difficulty label.

Surface and collision constraints anchor the abstract corridor in a usable
world.  They include elevation and grade when relevant, road-surface support,
boundary collision geometry, obstacles, lane markings whose semantics matter to
the task, spawn configuration, and an unambiguous world frame.  A course can be
geometrically sound in its source representation yet fail as a simulation
artifact because meshes, collision shapes, units, frames, or route identifiers
are lost on export.  OpenDRIVE-based construction demonstrates the useful but
limited role of interchange: a generated road network can be loaded by multiple
tools that derive meshes and waypoints, but import success is an interoperability
property, not evidence that the export operation generated the course
\cite{Muktadir2023ProceduralGeneration}.  The corresponding course card should
record source geometry, collision realization, semantic loss, and import/reset
outcomes separately, as required by \cref{sec:metrics,sec:benchmark-protocol}.

\subsection{Aerial Courses}

An aerial course is not a ground corridor lifted into three dimensions.  Its
task-defining markers are typically gates with a three-dimensional pose,
orientation, aperture, and crossing direction or precedence.  Position alone is
not enough: a gate's normal can determine which passage counts, its aperture
sets a spatial tolerance, and its orientation affects the approach trajectory and
visibility.  Drone racing is consequently a sequential navigation problem
through gates at speed, with collision-free flight and lap-time objectives
\cite{Kaufmann2018DeepDrone}.  The course contract should provide a total order,
partial order, or admissible transition relation for gates and define how a
crossing is detected, including tolerance, direction, and reset behavior.

Free space is the aerial analogue of a ground corridor but is generally more
coupled to vehicle dynamics.  Obstacles, keep-out volumes, gate-frame geometry,
minimum clearance, and the world boundary specify where flight is allowed.
Visibility is a separate constraint: a gate may be geometrically reachable while
occluded, out of sensor range, or visually confusable with an obstacle.  Recent
vision-based drone-racing work identifies the joint effect of coarse gate
position information, depth ambiguity between gates and obstacles, limited
feasible space, and the speed--collision-avoidance tradeoff
\cite{Yu2025MasteringDiverse}.  These are reasons to encode gate pose, obstacle
geometry, sensing assumptions, and occlusion-relevant realization together,
rather than describing a course only as a sequence of points.

Feasibility must also be dynamics conditioned.  Gate spacing, relative
orientation, turn radius, acceleration and braking margins, thrust limits,
attitude limits, and a vehicle's collision body determine whether a nominal gate
sequence is traversable.  A nominally collision-free center trajectory is not a
certificate for a vehicle with finite extent or a controller with bounded
tracking error.  Simulators such as Flightmare and Aerial Gym are valuable
supporting realizations for high-rate aerial evaluation, but a simulator API does
not establish the gate semantics or feasibility predicate for a particular
course \cite{Song2021FlightmareFlexible,Kulkarni2025AerialGym}.  The course
specification should therefore declare the vehicle/control envelope, collision
model, frames, units, and termination conditions in addition to spatial markers.

\subsection{Maritime Courses}

Maritime courses are organized around waypoints, buoys, marks, gates, regions,
and task rules rather than a continuously bounded paved corridor.  Their
semantics must say whether a mark is to be passed on a particular side, whether a
virtual or physical buoy must be reached, which gate direction is valid, and
whether the relation is a sequence, a choice, or a coverage task.  The World
Robotic Sailing Championship fleet-race rules, for example, define a start line,
ordered markers, a finish line, and completion assessed from tracked positions
\cite{WRSCIRSCOrganizingCommittee2019WorldRobotic}.  Such rules provide a
concrete specification of task semantics and scoring; they do not provide a
course-synthesis method.

Unlike a Cartesian indoor scene, a maritime realization often needs explicit
geospatial frames and transformations.  Course data can combine latitude and
longitude, local tangent-plane coordinates, compass headings, and simulator
world coordinates.  A portable record must identify the datum, projection or
local frame, units, heading convention, and numerical tolerance used to decide
passage or arrival.  The official sailing rules illustrate this need by defining
GPS-based tracking and SI-unit scoring alongside latitude/longitude conventions
\cite{WRSCIRSCOrganizingCommittee2019WorldRobotic}.  Without a declared frame,
a seemingly small coordinate conversion can move a virtual mark, reverse a
passage-side test, or make a simulator realization irreproducible.

Navigable-region and environmental predicates are equally important.  The course
may constrain water boundaries, shallow or excluded regions, obstacle and vessel
separation, current or wind conditions, time limits, right-of-way rules, and
communication or multi-vessel task conditions.  A waypoint search procedure can
mutate a pre-existing route while optimizing a testing objective, as in WPgen;
the altered waypoints remain meaningful only relative to a vessel model,
navigation software, and collision-aware simulation context
\cite{Nylnder2025SearchBased}.  RoboBoat and RobotX task documents similarly
specify mission, safety, and obstacle-avoidance conditions that a simulator or
generator must preserve when they are task defining
\cite{RoboNation2026RoboBoat2026,RoboNation20222022Maritime}.  Rules should be
encoded as predicates or evaluation conditions, with a clear statement of which
environmental quantities are fixed, sampled, observed, or left outside the
course contract.

\begin{table}[t]
  \centering
  \caption{Domain constraints that must be separated from an implementation's
  incidental rendering or storage choices.}
  \label{tab:domain-constraint-map}
  \small
  \begin{tabularx}{\linewidth}{@{}p{0.16\linewidth}Y Y Y@{}}
    \toprule
    Domain & Spatial support & Semantic markers and traversal relation & Feasibility and realization contract \\
    \midrule
    Ground & Surface-bound corridor, lanes, boundaries, grade, obstacles &
    Route/lane successors, checkpoints, start and finish & Width, curvature,
    clearance, collision geometry, vehicle envelope, units/frames, export and
    import behavior \\
    Aerial & Three-dimensional free space, obstacles, keep-out volumes & Gate
    pose, aperture, crossing direction, precedence & Clearance, visibility,
    dynamics and controller envelope, collision body, sensor assumptions,
    frames and reset behavior \\
    Maritime & Navigable water region, boundaries, hazards, environmental field &
    Waypoints, marks, buoys, gates, pass side, sequence or coverage relation &
    Geospatial frame, passage tolerance, vessel envelope, collision/right-of-way
    rules, weather/current assumptions, timing and scoring \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Shared Course Abstraction}

The domain differences above can be expressed through five common components:
a \emph{traversal relation} over task elements; \emph{spatial support} in
which traversal may occur; \emph{semantic markers} that identify gates,
boundaries, checkpoints, buoys, or regions; \emph{feasibility predicates}
evaluated under a declared vehicle, controller, and environment envelope; and a
\emph{realization contract} that preserves the relevant elements across
serialization and simulation.
This abstraction is intentionally not a claim that all domains share the same
geometry.  A ground course's support is normally a narrow surface corridor; an
aerial support is a volume; a maritime support is a navigable geospatial region
whose boundaries and dynamics may vary with conditions.  What transfers is the
need to make progression, allowed space, task markers, feasibility, and
realization explicit.

The abstraction also separates course constraints from fixed competition rules.
Rules are supporting constraints when they determine marker-passage semantics,
coordinate conventions, safety limits, scoring, or completion.  They remain
external to course generation unless a method explicitly transforms those
constraints into new spatial course structure.  Treating rules this way permits
benchmark and simulator interfaces to support reproducibility without inflating
them into generation evidence.  It also makes cross-domain comparison more
defensible: comparison can occur at the level of declared traversal relation and
predicate, while the domain extension retains the surface, volume, or maritime
environmental detail that determines whether a course is actually feasible.
