\section{Metrics and Reporting Protocol}
\label{sec:metrics}

This section proposes an auditable metric framework for course generators and their
downstream use. It is a reporting design, rather than a claim that the literature
already reports these measures or that a benchmark has been released. The framework
keeps four reporting units distinct: a \emph{generator} is a versioned sampling
procedure; a \emph{course} is one sampled, serialized course; a \emph{simulator
instance} is an imported and resettable realization of that course; and a
\emph{policy evaluation} is a specified policy, seed, course, and simulator run.
Aggregates should identify their unit and sampling population rather than mixing
these levels.

For a collection of $N$ generated courses, let $I_i^{(q)}\in\{0,1\}$ indicate that
course $i$ passes check $q$. The reported yield for that check is
\begin{equation}
  \widehat{p}_q = \frac{\sum_{i=1}^{N} I_i^{(q)}}{N},
  \label{eq:metric-yield}
\end{equation}
where the numerator is the number of generated courses passing $q$ and the
denominator is \emph{all} generation attempts, including rejected or repaired
ones. Report $N$, the number of independent generator seeds or runs, and a binomial
confidence interval (for example, a Wilson interval) or a seed-level bootstrap
interval. Continuous course and policy summaries should report their estimator,
number of courses, number of policy seeds where applicable, and a bootstrap interval
over the declared independent unit.
At the worst-case yield $p=0.5$, a 95\% binomial interval has approximate
half-width $0.98$ percentage points at $N=10{,}000$ and $0.31$ percentage points at
$N=100{,}000$. These are calibration values for an adaptive sampling budget, not
evidence that either count is universally sufficient.

\subsection{Feasibility}

Feasibility is a chain of declared checks, not a single validity label. At the
generator level, report completion yield: the fraction of calls that emit a course
artifact before timeout or resource limit. At the course level, separately report
geometric and topological validity, minimum boundary or obstacle clearance,
closed-loop status when a loop is required, and ordered-course status when an ordered
sequence of gates, waypoints, or route elements is required. For each predicate,
state its tolerance, coordinate frame, and the course population to which it applies.

Dynamics-aware feasibility adds a vehicle and control envelope $E$ to the contract.
For course $i$, define $I_i^{(\mathrm{dyn},E)}=1$ only if a declared model, initial
state set, control limits, and termination conditions admit a feasible traversal (or
the stated reachability proxy). This quantity is conditional on $E$; it must not be
presented as an intrinsic property of geometry. Report the solver or controller,
budget, tolerance, and inconclusive cases separately from failures.

At the simulator level, measure import, instantiation, reset, and collision-check
yields with the same numerator--denominator discipline as
\cref{eq:metric-yield}. A course that is geometrically valid but cannot be imported
is a simulator failure, not evidence of generator validity. Retain raw outcomes
before repair, and, when repair is offered, report repaired yield and repair rate as
additional quantities.

\subsection{Geometry}

Geometry descriptors characterize a valid or attempted course without calling it
better or harder. Report length and empirical distributions of curvature $\kappa(s)$
and curvature rate $d\kappa/ds$ over arc length $s$, using stated discretization and
units. Where applicable, also report slope, elevation range, and torsion; width and
clearance profiles; gate spacing, orientation, and visibility; and topology such as
component count, loop closure, crossing count, branching factor, and route-choice
structure. Use medians and quantiles alongside means for skewed descriptors, and
state whether non-finite values from invalid courses are excluded, separately counted,
or assigned a sentinel.

\subsection{Difficulty}

Difficulty is proposed as a continuous, multidimensional spectrum rather than the
labels ``easy,'' ``medium,'' and ``hard.'' Let $d_i\in\mathbb{R}^m$ be a declared
descriptor vector for course $i$, with coordinates selected from the geometric,
visibility, clearance, topology, and dynamics-aware quantities above. Normalize each
coordinate within a declared domain and reference distribution, for example
\begin{equation}
  z_{ij}=\frac{d_{ij}-\operatorname{median}(d_{\cdot j})}
  {\operatorname{IQR}(d_{\cdot j})},
  \label{eq:descriptor-normalization}
\end{equation}
with explicit handling when the interquartile range is zero. A benchmark may then
stratify by marginal or joint descriptor quantiles, coverage cells, or a continuous
curve through descriptor space. The selected strata and their boundaries are
reporting devices, not difficulty classes.

Comparative coverage uses a published, frozen reference population, descriptor
transform, bin edges, and absolute admissible descriptor bounds. Construct it from a
pooled pilot plus curated domain anchors independent of any one evaluated generator.
Generator-relative quantiles remain diagnostics only; they cannot establish
comparative coverage. Release reporting requires a fixed absolute coverage floor and
occupancy of every fixed spectrum stratum defined in
\cref{sec:benchmark-protocol}.

Policy-derived descriptors, such as success probability, required tracking bandwidth,
or margin to constraint violation, are reference-agent dependent. They should name the
reference policy, training distribution, policy seed set, simulator, and evaluation
budget. Robustness margins should be reported as perturbation radii or distributions
over stated initial-state, model, sensor, or actuation perturbations. A universal
scalar difficulty score is unsafe because it can conceal incompatible demands and can
invert when the vehicle, controller, objective, or normalization domain changes.

\subsection{Diversity}

Diversity is evaluated in descriptor space conditional on validity. Let
\begin{equation}
  \mathcal{V}=\{i\in\{1,\ldots,N\}:I_i^{(\mathrm{valid})}=1\},
  \qquad n_V=|\mathcal{V}|,
  \label{eq:valid-index-set}
\end{equation}
where the validity predicate is declared. Given normalized descriptors $z_i$, report
coverage of a declared grid, quantile lattice, or radius-based support on
$\mathcal{V}$; a duplicate or near-duplicate rate with its numerator, denominator,
and representation-level threshold evaluated on $\mathcal{V}$; and, for $n_V\geq2$,
the pairwise summary
\begin{equation}
  \widehat{D}_{\mathrm{pair}}
  =\binom{n_V}{2}^{-1}\sum_{\substack{i<j\\i,j\in\mathcal{V}}}
  \lVert z_i-z_j\rVert.
  \label{eq:feasible-pairwise-diversity}
\end{equation}
For collections $A$ and $B$, energy-distance or maximum-mean-discrepancy-style
estimators use their corresponding valid subsets $A_V$ and $B_V$ and report the
kernel or distance, normalization, sample sizes, and uncertainty. Entropy is
appropriate only when binning or density estimation is supported by the valid sample
size; report the estimator, bias correction if any, and sensitivity to binning or
bandwidth.

All such diversity estimates are feasible diversity and must be reported beside raw
generation yield. A summary over all attempts is permissible only when explicitly
labelled raw, with its invalid/missing-descriptor treatment and attempted-course
denominator stated. This prevents a collection from appearing diverse only because
invalid samples spread through descriptor space.

\subsection{Sim-to-Real, Recoverability, and Interaction}

Sim-to-real comparison uses a fixed real reference corpus, a versioned extraction
procedure, and declared conditioning variables such as vehicle, site, weather, or
mission. Compare the generated and real descriptor distributions with a stated
distance or test, sample sizes, and uncertainty; an unconditioned match does not
establish agreement within a deployment stratum.

Recoverability is conditional on a declared perturbation distribution, recovery
horizon, controller, and compute or action budget. Record whether the trajectory enters
a declared safe set, reaches a recovered set within the horizon, collides, or incurs an
irreversible failure. Report the denominator over perturbation draws and separate
inconclusive budget exhaustion from failure.

For multi-agent or overtaking courses, declare agent roles, interaction and pass rules,
minimum separation, and initial-state distribution. Report collision, deadlock, legal
pass, illegal-pass, and completion outcomes per interaction episode, with paired seeds
where policies are compared.

\subsection{Learned and Language-Model Generator Fidelity}

For learned or language-model generators, report offline data and compute, online
calls, tokens, and wall time, plus validation and repair cost; alternatively publish a
preregistered cost conversion. Evaluate four distinct fidelities: constraint adherence,
distributional agreement, semantic-intent agreement, and deterministic reconstruction.
Semantic explanations are tested against generator traces and declared constraints by
precision/recall or intervention consistency, not subjective prose.

\subsection{Reproducibility}

Each generator-level report should record generator identifier and version, complete
configuration, seed or seed schedule, dependency and simulator versions, vehicle and
control envelope, and a content hash for each emitted course and suite manifest.
Report determinism by reconstructing the same inputs in fresh processes and comparing
artifact hashes or a declared numerical tolerance. Runtime, throughput, peak memory,
and timeout rate require hardware, parallelism, warm-up policy, and reporting unit
(for example, seconds per attempted course and feasible courses per GPU-hour).
Finally, report rejection rate, repair rate, repair runtime, and the distribution of
attempts per retained course so that apparent output quality is not detached from its
cost.

\subsection{Simulation Feasibility}

Simulation feasibility, or interoperability, evaluates the path from portable course
representation to usable learning environment. At the simulator level, report
serialization and import success; declared units and coordinate frames; collision
geometry agreement with the source representation; spawn and reset success; and
deterministic reconstruction of a course from its manifest. Check sensor and world
integrity after import, including required observations, static geometry, and absence
of unintended collisions or missing assets.

The interface contract should state whether the simulator exposes a Gymnasium-style
reset/step API, supports vectorized environments, preserves course identifiers and
seeds in metadata, and runs headlessly. Report headless steps or episodes per second,
parallel environment count, hardware, and reset cost. These are implementation-target
measures: a portable representation does not by itself establish support in every
simulator.

\begin{table}[t]
  \centering
  \caption{Proposed auditable metric map. Every estimate identifies its reporting
  unit, denominator or sample population, and uncertainty procedure.}
  \label{tab:metric-map}
  \small
  \begin{tabularx}{\linewidth}{@{}Y Y Y@{}}
    \toprule
    Question & Metrics and units & Failure interpretation \\
    \midrule
    Does the generator emit valid courses? & Completion, geometric/topological,
    clearance, ordered/closed-loop, and dynamics-aware yields; passed attempts / all
    attempts & Generator failure, declared-constraint failure, or envelope-conditional
    infeasibility \\
    What courses does it produce? & Length, curvature, slope/elevation/torsion,
    width/clearance, gate, and topology distributions; native physical units & A
    descriptor is missing, ill-defined, or outside the declared domain \\
    Where is a course on the spectrum? & Normalized descriptor vector, quantile or
    coverage-cell assignment, robustness margin; dimensionless coordinates plus native
    margins & A scalar or categorical label hides dependence on domain or reference
    agent \\
    How varied is the feasible output? & Feasible coverage, pairwise distance,
    energy/MMD-style distance, entropy when estimable, near-duplicate rate & Diversity
    is low, estimated unreliably, or inflated by invalid samples \\
    Can another system reproduce and use it? & Manifest/hash match, runtime,
    memory, rejection/repair cost, import/reset/collision and headless yields & Missing
    provenance, nondeterminism, or a broken simulator interface \\
    Does generation agree with deployment data? & Fixed real-reference corpus,
    extraction/conditioning, distribution distance, and interval & Reference mismatch,
    insufficient conditioning, or uncertain estimate \\
    Can the system recover and interact safely? & Safe/recovered, collision,
    irreversible-failure, deadlock, and legal-pass rates & Recovery, interaction, or
    rule failure under the declared budget \\
    Is a learned generator faithful under budget? & Offline/online/repair cost and
    constraint, distributional, semantic, reconstruction fidelity & A selected output
    hides budget, trace, or reconstruction failure \\
    \bottomrule
  \end{tabularx}
\end{table}

\Cref{sec:benchmark-protocol} turns these metric contracts into a proposed sampling,
split, and evaluation protocol. In particular, it preserves raw feasibility outcomes
before descriptor-based selection and evaluates policies on common immutable course
identifiers through the paired, course-level comparisons specified in
\cref{sec:benchmark-protocol}.
