\section{Benchmark Protocol}
\label{sec:benchmark-protocol}

This section proposes a reference benchmark protocol built on the metric contracts in
\cref{sec:metrics}. It specifies release requirements separately from future
extensions. It neither claims that a released benchmark exists nor attributes these
choices to the current literature.

\subsection{Generator and Seed Contract}

Each participating generator should be exposed through a versioned adapter with a
stable identifier, declared input schema, complete configuration, and seed contract.
For each requested course identifier, the adapter records generator version,
configuration hash, root seed and derived seed, vehicle/domain assumptions, output
format version, and content hash of the raw artifact. The contract must state whether
the same inputs reconstruct byte-identical output or only output within a numerical
tolerance. A manifest should retain failed attempts, termination reason, wall-clock
budget, and any repair operation; repair may create a derived artifact, but must never
replace the raw feasibility record.

The proposed distribution-estimation and training target is $10{,}000$--$100{,}000$
attempted courses per generator, selected according to generator cost and domain
complexity. This range is a target for estimating descriptor distributions and
supporting training, not a universal adequacy claim. Reports should give the attempted
count, feasible count, raw and repaired yields, independent generator seeds, and
uncertainty as defined in \cref{sec:metrics}.

At the worst-case yield $p=0.5$, 95\% binomial half-width is about $1.0$ percentage
point for 10,000 attempts and $0.31$ points for 100,000. The range is therefore an
initial adaptive budget, not a universal mandate. Before sampling, pre-register
stopping criteria for yield-interval tolerance, descriptor quantile and coverage
stability, and a minimum valid count in every fixed spectrum stratum. Stop early only
when all criteria are met. If 100,000 attempts leave any criterion unmet, report
non-convergence or the shortfall rather than claiming comparable distributions.
\subsection{CourseSpec and Course Cards}

The normative portable IR is \texttt{CourseSpec}. It records geometry, topology,
gates, and waypoints; bounds, widths, and clearance; elevation; frames and units;
semantic obstacles and checkpoints; the domain and vehicle/control envelope; and
provenance, format version, and content hash. Raw-to-repaired lineage is explicit:
repair creates a derived artifact and never replaces the raw attempt or its failure
record. Domain extensions preserve constraints that the common core cannot express.

Each released course card contains immutable identifiers and hashes; generator,
configuration, and seed lineage; the reference distribution and descriptor
normalization; perturbation and recovery contract; agent interaction and pass rules;
units, frames, semantics, and semantic loss on export; conformance/import/reset status;
and repair, validity, and known-limitations records. The card is the course-level
interface between the metric definitions in \cref{sec:metrics} and the adapters below.

\subsection{Course Construction and Splits}

The construction order is fixed. First, generate raw attempts and evaluate feasibility.
Second, compute descriptors for the declared population, retaining explicit missing or
invalid status. Third, form spectrum strata and diversity-constrained selections from
the valid subset. This order ensures that repair and selection cannot erase raw
generator failure rates. Any repair budget, feasibility predicate, descriptor code,
normalization reference set, and selection algorithm belongs in the versioned suite
manifest.

For generator comparison, publish and freeze the reference population, descriptor
transform, bin edges, absolute admissible bounds, coverage floor, and fixed-stratum
occupancy before evaluation. Build the reference from a pooled pilot plus curated
domain anchors independent of any single generator. Generator-relative quantiles are
diagnostics only, not comparative coverage evidence.

An immutable held-out evaluation suite may admit a declared domain/configuration only
after its candidate pool contains at least 100 feasible, non-duplicate courses spanning
the declared continuous descriptor spectrum and diversity constraints, such as a
minimum separation or coverage target. These courses are never named ``easy,''
``medium,'' or ``hard.'' Instead, the manifest provides their descriptor vectors,
quantile memberships, normalization domain, and feasibility evidence. A shortfall is a
generation-audit result: it reports the candidate population and failure causes, but
does not authorize a smaller immutable released suite for that domain/configuration.

The 100-course threshold is a release floor for fixed-stratum spectrum coverage, not
a precision claim: its worst-case binary 95\% half-width is about 10 percentage
points. Expand the suite until preregistered precision or power is met. An infeasible
configuration remains audit-only; it cannot enter a smaller released suite.

Training, validation, and test partitions must separate root seeds and configuration
families. A configuration family includes all parameterizations that share a common
template or latent design source defined by the adapter. Before release, check for
cross-split duplicates and near duplicates using both content-hash identity and a
declared representation/descriptor threshold. Record the check, threshold, candidate
pairs, and removals in the suite manifest; seed separation alone is not a leakage
guarantee.

\subsection{Policy Evaluation}

For a global policy/training seed set $p\in\mathcal{P}$, train each seed once and
evaluate it on every course $c\in\mathcal{C}$; rollout seeds
$r\in\mathcal{R}_{pc}$ are nested only within the $(p,c)$ cell. Let $S_{cpr}$ and
$V_{cpr}$ be success and violation indicators. For example,
\begin{equation}
  \overline{S}_{cp}=\frac{1}{|\mathcal{R}_{cp}|}\sum_{r\in\mathcal{R}_{cp}}S_{cpr},
  \qquad \widehat{S}=\frac{1}{|\mathcal{C}||\mathcal{P}|}
  \sum_{c\in\mathcal{C}}\sum_{p\in\mathcal{P}}\overline{S}_{cp}.
  \label{eq:policy-success-estimator}
\end{equation}
Define $\overline{V}_{cp}$ and $\widehat{V}$ analogously. The estimand averages over
the declared course and global training-seed populations; its denominators are
rollouts per $(p,c)$ cell, global training seeds, and immutable-suite courses.

Report time in seconds (lap time or time to termination), progress as a declared
fraction in $[0,1]$ or distance in metres, position and angular tracking error in
metres and radians, energy in joules when defined, and otherwise-integrated control
effort in declared action-squared-seconds units. Each continuous outcome states its
denominator and aggregation order. A timeout without completion has $S_{cpr}=0$;
collisions and rule violations set $V_{cpr}=1$ under the declared taxonomy. Timeouts,
collisions, and other terminations remain distinct outcome labels, and time, progress,
tracking, and effort state whether failures are included, censored at the horizon, or
reported only for successful runs.

Evaluation is paired across common immutable course identifiers. For two policies,
form paired differences after rollout aggregation. Preserve method pairings with common
course IDs, training-seed schedule or index, and rollout random streams where
meaningful. Use hierarchical paired resampling: resample global training seeds as
clusters across all sampled courses, resample courses while preserving common IDs, then
resample rollout seeds within each $(p,c)$ cell. Report course sensitivity and global
training-seed sensitivity. An unmatched design requires an appropriate crossed
mixed-effects model rather than false pairing. Retain the failure taxonomy.
Training cost should include environment interactions, simulator steps, wall-clock
time, and courses observed, so sample efficiency is not conflated with throughput.
Cross-generator transfer evaluates held-out generators or configuration families;
cross-simulator transfer evaluates a portable course and policy contract in a second
declared simulator. Neither result is implied by in-domain success.

\subsection{Cross-Generator Training-Utility Protocol}

Let $G_1,\ldots,G_m$ denote participating generators. For every $G_i$, train the
same RL algorithm under one vehicle, dynamics, simulator, observation/action, reward,
termination, policy-architecture, optimization, environment-interaction, and
policy-update contract, using the common policy-seed set $P$. Only the distribution
of training courses changes across generator rows. Equal environment-interaction
budget is the primary training-budget comparison. Before training, freeze the
course-sampling and reuse contract. Its manifest reports the complete set of unique
training-course identifiers and content hashes, generator seeds, sampling
probabilities and presentation order, per-course reuse counts, curriculum or
scheduling rule, and all generation attempts and failures. Separately report
simulator steps, policy updates, wall-clock time, and compute so that equal
interaction budgets do not conceal generator or throughput cost.

Evaluate every resulting policy on the immutable suites $E_1,\ldots,E_m$ and on a
frozen evaluation-only real-track suite $E_{\mathrm{real}}$. Training on the finite
real-track corpus is excluded from this protocol because it would confound training
utility with memorization and augmentation policy. Preserve common course identifiers
and, where meaningful, common rollout random streams across training-generator rows.
Thus the crossed outcomes use the Section~\ref{sec:metrics} definition of $M^Y_{ij}$
and its real-track column; within a fixed evaluation column, compare training
distributions through \cref{eq:cross-generator-transfer-gap}, rather than treating
diagonal--off-diagonal differences over unlike course populations as the primary
transfer estimand.

Construct each generated suite feasibility-first: account for raw attempts and repair
outcomes before selection, then apply the same frozen descriptor transform, absolute
admissible bounds, spectrum strata, diversity rules, and minimum suite-size contract
to every generator. Before freezing the suites, perform manifest-backed duplicate and
near-duplicate checks between every realized training population and every generated
$E_j$ plus $E_{\mathrm{real}}$, using content-hash identity and the declared
representation/descriptor threshold. Report candidate pairs, decisions, removals,
and any resulting suite-size or stratum-occupancy shortfall rather than silently
replacing courses.

Require a suite-specific analysis over every course in each released $E_j$. Also
require a common-stratum analysis: before evaluation, freeze the intersection of
descriptor strata represented across all compared generators, fixed weights for
those strata, and each restricted per-column suite $E_j^{\cap}\subseteq E_j$ with
course set $C_j^{\cap}$. For that analysis, the $C_j$ entering the
Section~\ref{sec:metrics} definition of $M^Y_{ij}$ is $C_j^{\cap}$, and
stratum-specific cells are combined using the frozen weights. The suite-specific
analysis measures transfer over each released suite; the common-stratum analysis
reduces support and difficulty confounding. Report every missing stratum and
insufficient-course shortfall; do not silently shrink the declared intersection,
weights, or restricted suites. Condition $E_{\mathrm{real}}$ on the same domain,
vehicle, units, frames, descriptor extraction, and feasibility contract, and freeze
its composition and conditioning variables before policy evaluation.

On the same evaluation courses, run a declared model-based reference controller using
one generator-independent configuration. Freeze its model, objective, constraints,
horizon, sampling budget, controller-seed set, compute budget, and termination rules;
calibrate that configuration once on a declared set disjoint from all training and
evaluation suites. Preserve its seed and compute contracts across suites and
generators. First report a joint RL/reference-controller outcome table crossing
completion, collision, violation, timeout, and inconclusive budget exhaustion. For
every subsequent conditional metric or reference-controller gap, name the exact
course subset, RL rollout subset, reference-controller seed subset, and denominator.
The default subset for lap-time ratios and speed comparisons in
\cref{eq:reference-controller-gap} is courses completed by both methods. Failure by
the reference controller, including failure shared with RL, is inconclusive for
geometric or dynamic infeasibility.

For each outcome, aggregate rollouts within policy-seed/course cells and use
hierarchical paired resampling of policy seeds, common course identifiers, and nested
RL rollout seeds; for reference-controller gaps, also resample reference-controller
seeds within course. Report paired intervals for within-column comparisons and retain
the complete failure taxonomy with outcome-specific denominators and censoring rules.
Keep yield, descriptor coverage, feasible diversity, cost, interoperability, transfer
gaps, and reference-controller gaps as separate outcomes; do not collapse them into a
generator ranking.

\begin{table}[t]
  \centering
  \caption{Proposed benchmark release requirements and explicitly future extensions.}
  \label{tab:benchmark-requirements}
  \small
  \begin{tabularx}{\linewidth}{@{}p{0.24\linewidth}Y Y@{}}
    \toprule
    Component & Release requirement & Future extension \\
    \midrule
    Generator contract & Versioned adapter, configuration and seed contract, raw
    outcome and artifact hashes & Additional generators, domains, and vehicle/control
    envelopes \\
    Course collection & $10{,}000$--$100{,}000$ attempted courses per generator as
    the proposed target; feasibility-first descriptors and recorded repair & Larger
    distributions and adaptive sampling studies \\
    Evaluation suite & Admission only after $\geq100$ feasible non-duplicate courses
    span the declared spectrum and diversity constraints & Expanded suite sizes and
    longitudinal challenge sets \\
    Policy results & Common IDs, multiple policy seeds, crossed generator-to-suite
    outcomes, real-track evaluation, declared model-based reference-controller
    calibration, intervals, and failure taxonomy & Additional policy and
    reference-controller families \\
    Interoperability & Documented portable representation plus declared simulator/RL
    adapter behavior & More simulator backends, sensor suites, and hardware-in-the-loop
    validation \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Co-Evaluation and Course Cards}

Generator and policy results should be co-evaluated, not collapsed into one score. A
generator report gives raw and feasible yield, descriptor coverage, diversity, cost,
and interoperability; a policy report gives paired outcomes conditional on a declared
course suite and simulation contract. A policy's success rate therefore cannot be
used alone to rank generators, and a generator's geometric diversity cannot be used
alone to rank policies.

Each released evaluation course uses the \texttt{CourseSpec} and expanded course card
specified above. Its reconstruction, import/reset, collision, and semantic-loss records
are conformance evidence, not a substitute for a generator or policy result.

\subsection{Interoperability Targets}

The required single-environment contract is
\texttt{reset(seed, course\_id) -> (obs, info)} and
\texttt{step(action)} $\to$ \texttt{(obs, reward, terminated,}\
\texttt{truncated, info)}. Its schema
manifest declares observation/action keys, shapes, dtypes, bounds, units, and frames;
fixed $dt$ and action hold; and termination, truncation, and failure codes.

Use named independent seed streams and reconstruct each course deterministically from
its content hash. Declare state and sensor replay tolerances for each field and target
platform. The required release contract is this single-environment reset/step,
content-hash, and seed contract with headless execution.

Vectorized and tensor adapters are optional. When offered, they declare stable
environment ordering, per-environment seeds and course identifiers, autoreset
semantics, and final-observation handling; report their headless throughput. The exact
adapter set, file formats, simulator coverage, and optional interface behavior remain
implementation targets for a future release, not evidence of a released capability.
