\section{Open Reference Implementations}
\label{sec:reference-implementations}

An open reference release should support comparison across the taxonomy rather than
declare one generator to be best. Small, inspectable implementations expose
assumptions, failure modes, and engineering trade-offs; a single system risks making
its representation and repair choices an implicit benchmark definition. The proposed
release is therefore a common contract plus exemplars from distinct families, evaluated
under \cref{sec:metrics,sec:benchmark-protocol}. It is a release target, not a claim
about current code or that literature examples are interoperable software.

\subsection{Canonical Representation and Pipeline}

The portable intermediate representation (IR), here called \texttt{CourseSpec}, should
be domain aware without being simulator owned. It contains centerlines, segments,
gates, waypoints, and topology; widths, bounds, and clearance; elevation, local frames,
and units; semantic obstacles and checkpoints; and the declared domain and
vehicle/control envelope. Every raw or derived artifact carries provenance, format and
generator versions, and a content hash. Domain-specific fields are explicit extension
blocks, rather than silently approximating aerial gates as planar lanes or maritime
navigability as a road boundary.

The component path is
\begin{center}
\small
\texttt{GeneratorAdapter $\rightarrow$ raw CourseSpec $\rightarrow$ validators
$\rightarrow$ [declared repair/projection] $\rightarrow$ metrics/spectrum
$\rightarrow$ exporters $\rightarrow$ simulator/RL adapters}.
\end{center}
Validators emit predicate-level outcomes and diagnostics. Repair or projection is
optional and produces a separately hashed derived \texttt{CourseSpec}; it must never
replace the raw artifact or conceal a failed attempt. Metrics and spectrum assignment
use the declared valid population and reference normalization in
\cref{sec:metrics,sec:benchmark-protocol}.

\subsection{Generator and Interoperability Contract}

Each adapter accepts complete configuration and a root seed; emits a raw
\texttt{CourseSpec} or structured failure/diagnostic record; and states its deterministic
reconstruction guarantee. It supports batch and stream generation, with derived seeds
recorded. Repair is a visible downstream operation, never hidden in the API. A release
records version and configuration hashes, attempt status, runtime/resource counters,
and the raw artifact hash.

The release should include constructive spline or gate-sequence, stochastic procedural,
search or evolutionary, learned or environment-design, and repair/projection exemplars.
The exact set depends on the screened taxonomy and engineering validation; the labels
are selection criteria, not claims that every family is already released.

Export begins with portable geometry and semantics. Mappings to OpenDRIVE, Lanelet2,
SDF, USD, or simulator-native formats are offered only where semantically applicable.
The tested support matrix must be published rather than assumed: import, collision
geometry, reset, and identifier preservation are conformance outcomes. Each simulator
adapter implements the required Gymnasium-style single-environment contract in
\cref{sec:benchmark-protocol}; vectorized or tensor adapters are optional and declare
ordering, seed, autoreset, and final-observation semantics.

\subsection{Artifacts and Release Boundary}

The proposed pipeline generates a $10{,}000$--$100{,}000$ attempted-course corpus for
distribution estimation and training, preserving raw outcomes, manifests, course cards,
licenses, and provenance. Immutable spectrum-spanning evaluation suites are released
only after satisfying Section~\ref{sec:benchmark-protocol} admission criteria,
including at least 100 feasible non-duplicate courses for the declared
domain/configuration. A smaller shortfall is an audit result, not a suite substitute.

\begin{table}[t]
  \centering
  \caption{Proposed open reference-release contract. This survey makes no current-code
  implementation claim; planned releases and literature examples are distinct status
  categories.}
  \label{tab:reference-release-contract}
  \small
  \begin{tabularx}{\linewidth}{@{}p{0.25\linewidth}Y Y@{}}
    \toprule
    Surface & Required release contract & Optional extension or non-release status \\
    \midrule
    Representation & Versioned \texttt{CourseSpec}, units/frames, semantics,
    domain/vehicle envelope, provenance, and hashes & Domain extensions and exporters
    after semantic validation \\
    Generation & Config-and-seed adapter, raw result or diagnostic, deterministic
    reconstruction statement, batch/stream operation, and no hidden repair & Further
    taxonomy exemplars after screening and engineering validation \\
    Validation and metrics & Predicate outcomes, repair/projection lineage, and
    Section~\ref{sec:metrics} descriptors/spectrum metadata & New validators,
    policy-conditioned measures, and calibration studies \\
    Interoperability & Portable geometry plus tested single-environment conformance &
    Tested OpenDRIVE/Lanelet2/SDF/USD or simulator-native mappings; vector/tensor
    adapters where applicable \\
    Corpus and suites & Attempt manifests, cards, licenses, and suites admitted under
    Section~\ref{sec:benchmark-protocol} & Larger corpora, domains, and longitudinal
    suites \\
    Status disclosure & Current code, planned reference release, and literature
    examples labelled separately & No capability follows from a label, plan, or example \\
    \bottomrule
  \end{tabularx}
\end{table}
