\section{Reporting Practices in RL and Control}
\label{sec:reporting-practices}

Reports should make four claims independently auditable: what a generator attempted,
what a course is, what a simulator executed, and how a policy performed. A simulator
import failure is not a generator result, and a policy result conditional on one suite
is not a generator ranking. The feasibility, spectrum, and paired evaluation
definitions are those of \cref{sec:metrics,sec:benchmark-protocol}; visualizations
complement, but do not replace, the relevant counts, distributions, and intervals.

\begin{table}[t]
  \centering
  \caption{Minimum reporting checklist. A record means a machine-readable artifact
  plus a concise human-readable card; proposed artifacts are labelled as such.}
  \label{tab:reporting-checklist}
  \small
  \begin{tabularx}{\linewidth}{@{}p{0.22\linewidth}Y Y@{}}
    \toprule
    Record & Minimum disclosure & Claim it supports \\
    \midrule
    Generator card & Identifier/version/hash; complete configuration and all seeds;
    domain/vehicle envelope; raw attempts, failures, rejections, diagnostics, validity
    definition, repair policy/before-after outcomes, runtime, compute, and peak memory &
    Generation yield, cost, and constraints \\
    Course card & Immutable identifier and raw/derived hashes; lineage, geometry,
    semantics, units/frames, topology, descriptor vector, spectrum reference,
    diversity-conditioning/selection, license, and provenance & Course-level
    feasibility, distribution, and reuse \\
    Simulator conformance record & Import/export version and outcome, unsupported
    fields and failed imports, frame/unit mapping, collision/reset checks,
    observation/action schema, $dt$, hardware, and adapter seed behavior & Executed
    simulator semantics and interoperability \\
    Policy evaluation record & Algorithm, version, and configuration; every training and
    rollout seed; suite/simulator hashes; train/validation/test split and leakage
    checks; outcome taxonomy, denominators, uncertainty, interactions, simulator
    steps, wall-clock time, and memory & Conditional policy performance and cost \\
    \bottomrule
  \end{tabularx}
\end{table}

The attempt log retains every requested generation, including invalid outputs, timeouts,
negative imports, and rejected candidates. State whether each statistic is over all
attempts, raw valid courses, or repaired valid courses, with its denominator. Courses
must not be selected after inspecting policy outcomes; for diversity conditioning or
spectrum selection, record the candidate population, reference set, thresholds,
removals, and resulting split membership. No cherry-picked tracks are presented as a
distributional result.

Reproducibility requires generator, validator, repair, descriptor, exporter, simulator,
and policy versions; complete configurations; root and derived seeds; course and
manifest hashes; units and coordinate frames; and hardware, parallelism, warm-up,
timeout, and measurement policies. Archive the cards, conformance records, policy
records, source or executable environment, licenses, and provenance. State whether each
artifact is released, proposed, or unavailable, including access restrictions. This
disclosure boundary prevents a planned benchmark or literature example from being
reported as a reproducible implementation.
