\section{Falsifiable Research Hypotheses}
\label{sec:open-problems}

These are falsifiable hypotheses, not evidence-ranked gaps. No hypothesis is
evidence-ranked until Pass 2 is complete. Before data collection or evaluation, each
future study must preregister descriptor bounds, any cost conversion, endpoints,
rejection criteria, and its analysis plan. Each comparison uses the measures in
\cref{sec:metrics} and immutable identifiers, lineage, and adapters from
\cref{sec:benchmark-protocol}.

\paragraph{H1: Cross-domain representation.}
Compare source and imported predicates for ground, aerial, and maritime
\texttt{CourseSpec} instances. Reject common-core adequacy when a required
domain constraint is lost. Required artifacts are source/derived IRs, extension blocks,
export traces, and predicate-level conformance records.

\paragraph{H2: Geometry versus control feasibility.}
Estimate the matched-stratum difference between geometry/topology yield and
dynamics/control-aware yield for a declared envelope. Reject geometry-only adequacy
when the pre-specified gap bound is exceeded. Required artifacts are vehicle/controller
versions, budgets, initial states, solver outcomes, and raw attempts.

\paragraph{H3: Difficulty calibration.}
Compare descriptor-stratum ordering across reference agents, seeds, and simulators.
Reject policy-independent calibration when rank stability falls below its pre-specified
bound. Required artifacts are frozen reference distributions, descriptor code, policies,
and paired outcomes.

\paragraph{H4: Representative versus adversarial generation.}
Compare held-out distribution distance, coverage, duplicates, and policy failures under
matched budgets. Reject representative-generation claims when adversarial improvement
violates fixed distributional bounds. Required artifacts are target/reference corpora,
candidate pools, selection logs, and policy evaluations.

\paragraph{H5: Co-adaptive UED versus repair neutrality.}
Separate the effect of a co-adaptive UED or curriculum from the neutrality of a repair
pipeline. For co-adaptive versus frozen-generator training, match representation,
validity processing, exposure budget, policy seeds, and reuse schedule; evaluate on
common immutable suites. Reject the co-adaptation claim when preregistered common-suite
outcome bounds fail.

For repair neutrality, retain every raw attempt and repair eligibility, failure, and
attrition. Estimate descriptor and downstream shifts on a preregistered paired subset
feasible both before and after repair, while separately reporting unpaired feasibility
changes. Reject repair neutrality when preregistered paired descriptor or downstream-
outcome shift bounds fail, or when unpaired feasibility or attrition bounds fail.
Required artifacts are teacher/curriculum snapshot identifiers, policy checkpoints,
random-stream schedules, exposure logs, raw-derived lineage, the pairing rule, and
outcome denominators.

\paragraph{H6: Cross-generator training utility and real-track transfer.}
Train matched policy seeds separately on each generator and evaluate the resulting
policies on the full crossed generator-to-suite matrix and an evaluation-only
real-track suite. Compare preregistered suite-specific gaps $\Delta^Y_{ij}$,
stratum-weighted gaps $\Delta^{Y,\cap}_{ij}$, and paired real-track contrasts
$\Delta^Y_{ik,\mathrm{real}}$. Predefine primary outcomes, directions, bounds, and
their multiplicity family. Use a frozen model-based reference controller to calibrate
policy failure without treating it as an oracle. Reject a generator training-utility
claim if any designated primary contrast breaches its bound under the preregistered
multiplicity rule, or if the suite-specific and common-stratum analyses yield opposite
bound decisions.
Required artifacts are training distributions and exposure logs, immutable generated
and real suites, descriptor strata, fixed stratum weights
and shortfall records, policy, rollout, and reference-controller seeds,
reference-controller configuration, paired outcomes, and failure labels.

\paragraph{H7: Recoverability.}
Compare recovery probability across perturbation distributions, controllers, and
budgets. Reject a recovery claim when safe-set or recovered-set occupancy misses its
bound, or collision/irreversible failure exceeds its ceiling. Required artifacts are
perturbation draws, horizons, safe/recovered definitions, trajectories, and failure
taxonomy.

\paragraph{H8: Multi-agent and overtaking behavior.}
Compare collision, deadlock, legal-pass, illegal-pass, and completion rates across
roles and interaction rules. Reject a passing or interaction claim when separation or
legal-pass bounds fail. Required artifacts are role assignments, initial states, pass
rules, interaction traces, and outcome labels.

\paragraph{H9: Card, reconstruction, and audit agreement.}
Compare course-card declarations with reconstruction and conformance results, then
measure agreement between pre-specified audit codes and their cited locators. Reject
completeness when a required card field, reconstruction, conformance predicate, or
audit agreement threshold fails. Required artifacts are cards, hashes, manifests,
replay logs, and coded ledgers.

\paragraph{H10: Learned and language-model fidelity.}
Compare learned or language-model generators with declared baselines under an equal
budget vector: offline data and compute; online calls, tokens, and wall time; and
validation/repair cost, or a pre-specified conversion. Estimate constraint,
distributional, semantic-intent, and deterministic-reconstruction fidelity. Reject a
fidelity claim when any pre-specified dimension fails. Test explanation and semantic
fidelity against generator traces and declared constraints using precision/recall or
intervention consistency, not subjective prose. Required artifacts are prompts,
traces, configurations, outputs, budgets, validators, and reconstruction tests.

\paragraph{H11: Maritime, 3D, and benchmark drift.}
Compare source-to-simulator conformance and policy outcomes for maritime navigability,
currents/wind, 3D gate visibility, and ordered traversal; rerun a fixed panel after
every version change. Reject coverage or maintenance claims on missing required
predicates, failed round trips, or unexplained version shifts. Required artifacts are
domain contracts, simulator mappings, versioned suites, compatibility records, and
deprecation logs.
