\section{Conclusion}
\label{sec:conclusion}

Course generation is part of the experimental method of robot racing. A course
determines more than where a robot moves: together with a vehicle envelope and a
simulator realization, it determines which sensing, planning, and control demands are
presented to a system. Treating tracks, gates, waypoints, and buoy courses as
inspectable artifacts therefore clarifies claims that are otherwise easy to conflate.
Generator completion, geometric validity, dynamics-aware feasibility, simulator
conformance, and policy success are different outcomes with different denominators.
Likewise, a representation, its sampled realization, an exported scene, and a
trajectory through that scene are not interchangeable objects.

The minimum reporting recommendation follows directly from these distinctions. A
generator result should disclose its version, configuration, seeds, vehicle and domain
assumptions, every generation attempt, failure and repair outcome, validity predicates,
runtime and compute, and immutable artifact identities. A course record should expose
geometry, topology, semantics, units, frames, raw-to-derived lineage, descriptor
values, and export or import loss. A simulator result should report reconstruction,
collision-geometry, spawn, reset, observation, and headless-execution checks. A policy
result should identify the immutable suite, simulator and policy versions, training
and rollout seeds, outcome taxonomy, denominators, and uncertainty procedure. These
records keep attractive retained samples from concealing generator rejection and keep
controller outcomes from standing in for course quality.

The corresponding benchmark recommendation is to generate distributions before
selecting suites, preserve raw feasibility outcomes before repair, and select held-out
courses through fixed descriptor and diversity criteria. Difficulty should be reported
as a continuous, multidimensional spectrum rather than with the labels \emph{easy},
\emph{medium}, or \emph{hard}. Curvature, clearance, visibility, topology, elevation,
environmental loads, and dynamics-aware margins can impose distinct demands, and their
ordering can change with vehicle, controller, and simulator. The proposed
$10{,}000$--$100{,}000$ attempts per generator are an adaptive starting range for
distribution estimation and training, not a universal sufficiency threshold. The
proposed 100-course evaluation requirement is an admission floor for feasible,
non-duplicate, spectrum-spanning candidates, not a precision guarantee. Immutable
identifiers and paired, course-level evaluation are then needed to compare policies
without discarding failure modes or inventing false pairings.

The evidence state limits what can presently be concluded. Corrected-packet duplicate
reratings are complete, but final Pass~1 projection still requires governed integration
into a sealed final-version primary or adjudication snapshot, final adjudication, and
accountable-author verification. Pass~2 contribution coding underwent an initial
reliability assessment and a prospective CODEBOOK-v2 assessment on a deterministic
fresh holdout; neither met the prespecified all-field gate. Quantitative taxonomy,
prevalence, and comparative claims are therefore frozen. The representation,
generation, and domain sections provide a source-backed qualitative synthesis, not a
frequency analysis of the retained literature. The current \TrackGen{} audit likewise
documents a narrow planar implementation surface and release gaps; it does not
establish that \texttt{CourseSpec}, simulator/RL adapters, large corpora, or benchmark
suites already exist.

The next empirical step is not another round of label-specific exceptions. It is a
structural redesign of the contribution taxonomy, followed by fresh independent
validation, governed integration and final adjudication of the completed corrected-packet
reratings, and accountable-author verification. The next engineering step is to implement small, validated
reference generators across the revised taxonomy; serialize their raw and derived
outputs through \texttt{CourseSpec}; exercise conformance in mainstream simulator and
reinforcement-learning interfaces; and then generate large training distributions and
immutable evaluation suites under the proposed admission rules. That sequence turns
the survey's central argument into a testable program: courses should be generated,
measured, transported, and audited with the same care applied to the learning and
control algorithms evaluated on them.
